### 2024-07-25 docs/SOSP 2024/rebuttal.txt

Q1: How does SilvanForge’s decision-forest-based ‘schedule space’ and scheduling language differ from prior languages for other domains, e.g., TVM’s? For instance, what operators are new to SilvanForge?

Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees. One of the paper's key contributions lies in recognizing the existence of a scheduling space for a domain as diverse as ours and demonstrating the significant performance variations across different schedules. The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like walk interleaving, walk unrolling, reductions etc. These cannot be supported with existing scheduling languages. {Can we show an example.}
This has parallels to how TVM extends Halide to tensor specific primitives like tensorization and cooperative data loading. Despite the fact that TVM and Halide operate on similar domains (deep learning vs. image processing), such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding the required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work has proposed a small fixed set of optimization techniques and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

### 2024-07-25 docs/SOSP 2024/rebuttal.txt

Q1: How does SilvanForge’s decision-forest-based ‘schedule space’ and scheduling language differ from prior languages for other domains, e.g., TVM’s? For instance, what operators are new to SilvanForge?

Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees. One of the paper's key contributions lies in recognizing the existence of a scheduling space for a domain as diverse as ours and demonstrating the significant performance variations across different schedules. The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like walk interleaving, walk unrolling, reductions etc. These cannot be supported with existing scheduling languages. {Can we show an example.}
This has parallels to how TVM extends Halide to tensor specific primitives like tensorization and cooperative data loading. Despite the fact that TVM and Halide operate on similar domains (deep learning vs. image processing), such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding the required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work has proposed a small fixed set of optimization techniques and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

Q2: Which SilvanForge optimizations are least likely to be supported by PyTorch (in the context of a decision forest compiled by modern Hummingbird), and why?
A: 
Hummingbird represents the decision tree inference computation as a PyTorch graph using tensor operations in order to take advantage of systems like PyTorch and TVM. While this enables GPU compilation and re-targetability, the tensor abstraction is not sufficient to express rewrites that SilvanForge performs. 
Hummingbird exposes all forms of parallelism to the tensor compiler. The generated code performs all tree walks in parallel, stores the results of individual tree walks in a tensor and finally performs a reduction. Flattening out this parallelism correctly into high-performance kernels is non-trivial and likely beyond the ability of current tensor compilers. While Hummingbird does perform a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks), it is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.

Representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations. We expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory.

Additionally, the generated tensor code relies on gathering individual features when required. It is very unlikely that the underlying tensor compiler will decide to read input rows into shared memory as SilvanForge does. The analysis of memory access and reuse is complicated due to the indirect accesses of trees and rows.  
SilvanForge gets around this with domain specific abstractions and knowledge.

* As concrete examples, it would be nearly impossible (without domain specific knowledge) to restructure the computation to correctly determine the optimal number of threads to use for tree and row parallelism, or to determine whether or not to interleave walks and to what depth.

* We find that in some cases, padding trees is very detrimental to performance (almost 2x in some cases). HB currently always pads trees to the same depth. SilvanForge can generate code that does not pad trees. Additionally, even when trees are padded, HB forces all trees in the model to have the same depth while this is not the case with SilvanForge.

### 2024-07-25 docs/SOSP 2024/rebuttal.txt

Q1: How does SilvanForge’s decision-forest-based ‘schedule space’ and scheduling language differ from prior languages for other domains, e.g., TVM’s? For instance, what operators are new to SilvanForge?

Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees. One of the paper's key contributions lies in recognizing the existence of a scheduling space for a domain as diverse as ours and demonstrating the significant performance variations across different schedules. The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like walk interleaving, walk unrolling, reductions etc. These cannot be supported with existing scheduling languages. {Can we show an example.}
This has parallels to how TVM extends Halide to tensor specific primitives like tensorization and cooperative data loading. Despite the fact that TVM and Halide operate on similar domains (deep learning vs. image processing), such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding the required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work has proposed a small fixed set of optimization techniques and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

Q2: Which SilvanForge optimizations are least likely to be supported by PyTorch (in the context of a decision forest compiled by modern Hummingbird), and why?
A: 
Hummingbird utilizes PyTorch by translating decision tree inference computation to a PyTorch graph. While this enables GPU compilation and re-targetability, as we discuss below the tensor abstraction is not sufficient to express rewrites that SilvanForge performs. 
The Hummingbird translation exposes all forms of parallelism to the tensor compiler. The generated code performs all tree walks in parallel, stores the results of individual tree walks in a tensor and finally performs a reduction. In the process Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks). This as discussed in our paper is not always useful (say when?). It is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.

Further, representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations either. Flattening out the different sources of parallelism correctly into high-performance kernels is non-trivial and likely beyond the ability of current tensor compilers (this line is not convincing). 
Note that the generated tensor code relies on gathering individual features when required. With this representation the analysis of memory access and reuse is complicated due to the indirect accesses of trees and rows. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does.   
FInally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory.

SilvanForge gets around all these with domain specific abstractions and knowledge.

### 2024-07-26 docs/SOSP 2024/rebuttal.txt

Q1: How does SilvanForge’s decision-forest-based ‘schedule space’ and scheduling language differ from prior languages for other domains, e.g., TVM’s? For instance, what operators are new to SilvanForge?

Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees. One of the paper's key contributions lies in recognizing the existence of a scheduling space for a domain as diverse as ours and demonstrating the significant performance variations across different schedules. The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like walk interleaving, walk unrolling, reductions etc. These cannot be supported with existing scheduling languages. {Can we show an example.}
This has parallels to how TVM extends Halide with DNN compilation specific primitives like tensorization and cooperative data loading. Despite the fact that TVM and Halide operate on similar domains (deep learning vs. image processing), such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.

@@

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding the required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work has proposed a small fixed set of optimization techniques and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

Q2: Which SilvanForge optimizations are least likely to be supported by PyTorch (in the context of a decision forest compiled by modern Hummingbird), and why?
A: 
Hummingbird utilizes PyTorch by translating decision tree inference computation to a PyTorch graph. While this enables GPU compilation and re-targetability, as we discuss below the tensor abstraction is not sufficient to express rewrites that SilvanForge performs. 
The Hummingbird translation exposes all forms of parallelism to the tensor compiler. The generated code performs all tree walks in parallel, stores the results of individual tree walks in a tensor and finally performs a reduction. In the process Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks). This as discussed in our paper is not always useful (say when?). It is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.

Representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations either. More specifically, the generated tensor code relies on gathering individual features when required. With this representation the analysis of memory access and reuse is complicated due to the indirect accesses of trees and rows. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does. Second, to the best of our knowledge, tensor compilers cannot perform interleaving as SilvanForge does (this is not part of their schedule space). Third, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory (and not to shared memory as in the case of SilvanForge). 
Finally, it would be nearly impossible to restructure the computation to correctly determine the optimal number of threads to use for tree and row parallelism since no template schedules can be specified for the kernels finally generated by the tensor compiler. SilvanForge addresses these problems with domain specific abstractions and knowledge.

The above discussion brings up a more practical system level problem ... 
* Systems challenges 
    -- no access to lower level details like shared memory, thread synchronization, etc at the tensor abstraction
    -- kernels are generated dynamically from the tensor graph and there is no scope for controlling how these kernels are optimized at the level of the tensor graph. TVM scheduling language is lower level.

Another aspect where SilvanForge and Hummingbird differ is support for different representations of the model. SilvanForge can generate code for different representations of the model (array, sparse, reorg) while Hummingbird only supports a single representation. As noted in the paper, different representations can have different performance characteristics and SilvanForge can tune the code to use the representation that is best suited for the model being compiled.

These views are partially validated by prior experiments on CPUs where a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals a massive performance gap. It is also not possible to express the various GPU specific configurations that SilvanForge can express, like parallelism on both tree and row, caching and reduction, with just tensor abstractions.

### 2024-07-26 docs/SOSP 2024/rebuttal.txt

Q1: Novelty of scheduling language.

Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees. One of the paper's key contributions lies in recognizing the existence of a scheduling space for a domain as diverse as ours and demonstrating the significant performance variations across different schedules. The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like walk interleaving, walk unrolling, reductions etc. These cannot be supported with existing scheduling languages. {Can we show an example.} This has parallels to how TVM extends Halide with domain-specific primitives for DNN compilation like tensorization and cooperative data loading. Despite the fact that TVM and Halide operate on similar domains (deep learning vs. image processing), such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.

@@

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding the required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work has proposed a small fixed set of optimization techniques and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

Q2: Comparison with Hummingbird

Hummingbird utilizes PyTorch by translating decision tree inference computation to a PyTorch graph. While this enables GPU compilation and re-targetability, as we discuss below the tensor abstraction is not sufficient to express rewrites that SilvanForge performs. The Hummingbird translation exposes all forms of parallelism to the tensor compiler. The generated code performs all tree walks in parallel, stores the results of individual tree walks in a tensor and finally performs a reduction. In the process Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks). This as discussed in our paper is not always useful (say when?). It is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.

@@

Representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations either. More specifically, the generated tensor code relies on gathering individual features when required. With this representation the analysis of memory access and reuse is complicated due to the indirect accesses of trees and rows. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does. Second, to the best of our knowledge, tensor compilers cannot perform interleaving as SilvanForge does (this is not part of their schedule space). Third, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory (and not to shared memory as in the case of SilvanForge). Finally, it would be nearly impossible to restructure the computation to correctly determine the optimal number of threads to use for tree and row parallelism since no template schedules can be specified for the kernels finally generated by the tensor compiler. SilvanForge addresses these problems with domain specific abstractions and knowledge.

The above discussion brings up a more practical system level problem. Expressing the tree inference problem as a PyTorch graph means that the decision tree compiler has no access to lower level details like shared memory, thread synchronization, etc. Kernels are generated dynamically from the tensor graph and there is no scope for controlling how these kernels are optimized at the level of the tensor graph. For example, there is no way to specify a TVM schedule on the tensor graph.

@@

Another aspect where SilvanForge and Hummingbird differ is support for different representations of the model. SilvanForge can generate code for different representations of the model (array, sparse, reorg) while Hummingbird only supports a single representation. As noted in the paper, different representations can have different performance characteristics and SilvanForge can tune the code to use the representation that is best suited for the model being compiled.

These views are partially validated by prior experiments on CPUs where a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals a massive performance gap. It is also not possible to express the various GPU specific configurations that SilvanForge can express, like parallelism on both tree and row, caching and reduction, with just tensor abstractions.

Rev#B: Are your compiler changes 'portable' across platforms?
The scheduling language, high-level IR and mid-level IR are fully shared. As described in the paper, except for a few target specific primitives, the scheduling language is common for both target processors. The same scheduling language is used to specify the structure of inference code on the CPU. When a schedule doesn't mark any loops as targetting a GPU dimension, CPU code is generated. The high-level and mid-level IR and optimizations on them are fully shared between the CPU and GPU compilation pipelines. The inference computation is represented as loops in both these IRs and therefore the, the representations and transformations on them are common. The compilation pipelines diverge starting starting at the lowering of MIR to low-level IR (as described in the paper). Even so, we are able to share significant amount of code between the CPU and GPU lowering due to how in-memory representations are abstracted in SilvanForge (Section 6). We found that a search heuristic is not needed on the CPU. Exploring only a few configurations is sufficient to find the best schedule. For the CPUs we tested on, we found that applying all optimizations is always beneficial and only a few configurations of the loop-nest needs to be explored.

Rev#C: Have people moved on from decision tree models?
As we establish in Section 1, decision trees are still widely used in practice and are likely to remain so for the foreseeable future. Also, the fact that their popularity has not declined over the last few years is evident from the last few Kaggle surveys. We believe that the retargetable performance that SilvanForge provides will be valuable to the large number of practitioners who use decision trees.

Rev#C: Support for reductions.
As mentioned in the paper, the parallel reduction support in MLIR currently is either for value types (scf.reduce) or for tensors (linalg). Both of these do not fit the requirements of generating code for decision tree inference. With scf.reduce, every accumulation creates a logical copy and eliminating these copies correctly is a non-trivial problem. Additionally, MLIR does not have a lowering path from scf.reduce to a GPU implementation. With linalg, the entire tensor of values that are to be reduced need to be written to memory first thus making it impossible to fuse the reduction into the inner-most loops.

@@

The reduction support we implement in SilvanForge is fairly general. Only the process of identifying reduction loops (Section 5) is specific to decision tree inference (tree loops are identified as reduction loops). All lowering of the ops in the reduction dialect are general and can be used for any other application that requires reductions. \TODO{Should we say that we can replace the reduction loop identification with a more general mechanism?}

Rev#E: Other metrics besides kernel time speedup. 
While we did not evaluate such metrics, it should not be difficult to optimize for them using SilvanForge. We believe that the generality of the scheduling language encapsulates a vast enough space and finding variants that are optimized for other metrics should be possible.

### 2024-07-26 docs/SOSP 2024/rebuttal.txt

Q1: Novelty of scheduling language.

Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees. One of the paper's key contributions lies in recognizing the existence of a scheduling space for a domain as diverse as ours and demonstrating the significant performance variations across different schedules. The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like walk interleaving, walk unrolling, reductions etc. These cannot be supported with existing scheduling languages. This has parallels to how TVM extends Halide with domain-specific primitives for DNN compilation like tensorization and cooperative data loading. Despite the fact that TVM and Halide operate on similar domains (deep learning vs. image processing), such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.

@@

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding the required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work has proposed a small fixed set of optimization techniques and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

Q2: Comparison with Hummingbird

Hummingbird utilizes PyTorch by translating decision tree inference computation to a PyTorch graph. While this enables GPU compilation and re-targetability, as we discuss below the tensor abstraction is not sufficient to express rewrites that SilvanForge performs. The Hummingbird translation exposes all forms of parallelism to the tensor compiler. The generated code performs all tree walks in parallel, stores the results of individual tree walks in a tensor and finally performs a reduction. In the process Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks). This as discussed in our paper is not always useful (Section 2). It is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.

Representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations either. Consider the example of walk-interleaving. SilvanForge’s tree-walk interleaving relies on unrolling only the trees whose walks are being interleaved to the same depth. It also reorders trees so that trees with the same depth are processed together. While it maybe possible to extend tensor compilers to perform interleaving, it will be very hard to mimic what SilvanForge does. Current tensor compilers optimize dense matrix computation and do not have the optimizations necessary to interleave instructions across iterations. Even if they did, picking the right tree-walks (same tree different rows, different trees same row etc.) to interleave would be a non-trivial problem.

The tensor compiler cannot perform the required caching and reduction optimizations either. The generated tensor code relies on gathering individual features when required. With this representation the analysis of memory access and reuse is complicated due to the indirect accesses of trees and rows. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does. Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate.

@@

The above discussion brings up a more practical system level problem. Expressing the tree inference problem as a PyTorch graph means that the decision tree compiler has no access to lower level details like shared memory, thread synchronization, etc. Kernels are generated dynamically from the tensor graph and there is no scope for controlling how these kernels are optimized at the level of the tensor graph. For example, there is no way to specify a TVM schedule on the tensor graph.

Another aspect where SilvanForge and Hummingbird differ is support for different representations of the model. SilvanForge can generate code for different representations of the model (array, sparse, reorg) while Hummingbird only supports a single representation. As noted in the paper, different representations can have different performance characteristics and SilvanForge can tune the code to use the representation that is best suited for the model being compiled.

These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals a massive performance gap.

### 2024-07-26 docs/SOSP 2024/rebuttal.txt

Q1: Novelty of scheduling language.

Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees. One of the paper's key contributions lies in recognizing the existence of a scheduling space for a domain as diverse as ours and demonstrating the significant performance variations across different schedules. The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like tree-walk interleaving, tree-walk unrolling, reductions etc. These cannot be supported with existing scheduling languages. This has parallels to how TVM extends Halide with domain-specific primitives for DNN compilation like tensorization and cooperative data loading. Despite the fact that TVM and Halide operate on similar domains (deep learning vs. image processing), such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding the required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work has proposed a small fixed set of optimization techniques, and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

@@

Q2: Comparison with Hummingbird

Hummingbird translates the decision tree inference into a series of tensor operations to use PyTorch. While this enables re-targetability, the tensor abstraction is insufficient to express all SilvanForge optimizations. Hummingbird exposes all forms of parallelism to the tensor compiler. The generated tensor code performs all tree walks in parallel, stores the individual tree predictions in a tensor and finally performs a reduction. In the process, Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks). This, as discussed in our paper, is not always useful (Section 2). It is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.

Representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it infeasible for the tensor compiler to perform these optimizations. Consider the example of walk-interleaving. SilvanForge’s tree-walk interleaving relies on unrolling only the trees whose walks are being interleaved to the same depth. It also reorders trees so that trees with the same depth are processed together. While it maybe possible to extend tensor compilers to perform interleaving, it will be very hard to mimic what SilvanForge does. Current tensor compilers optimize dense matrix computation and do not have the optimizations necessary to interleave instructions across iterations. Even if they did, picking the right tree-walks (same tree different rows, different trees same row etc.) to interleave would be a non-trivial problem.

The tensor compiler cannot perform the required caching and reduction optimizations either. Since the generated tensor code relies on gathering individual features when required, the indirect accesses of trees and rows complicates the analysis of memory access and reuse. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does. Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.

The above discussion brings up a more practical system level problem. Expressing the tree inference using PyTorch means that the decision tree compiler has no control over lower level details like shared memory, thread synchronization, etc. Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized. Systems like TorchDynamo and TorchInductor do not provide this level of control to the PyTorch program.

These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap.

Rev#B: Are your compiler changes 'portable' across platforms?
The scheduling language, high-level IR and mid-level IR are fully shared. The same scheduling language is used to specify the structure of inference code on CPUs and GPUs. When a schedule doesn't mark any loops as targetting a GPU dimension, CPU code is generated. The high-level and mid-level IR and optimizations on them are fully shared between the CPU and GPU compilation pipelines. The inference computation is represented as loops in both these IRs and therefore the, the representations and transformations on them are common. The compilation pipelines diverge starting starting at the lowering of MIR to low-level IR. Even so, we are able to share significant amount of code between the CPU and GPU lowering due to how in-memory representations are abstracted in SilvanForge (Section 6). We found that a search heuristic is not needed on the CPU. Exploring only a few configurations is sufficient to find the best schedule. For the CPUs we tested on, we found that applying all optimizations is always beneficial and only a few configurations of the loop-nest needs to be explored.

### 2024-07-26 docs/SOSP 2024/rebuttal.txt

The tensor compiler cannot perform the required caching and reduction optimizations either. Since the generated tensor code relies on gathering individual features when required, the indirect accesses of trees and rows complicates the analysis of memory access and reuse. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does. Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.

The above discussion brings up a more practical system level problem. Expressing the tree inference using PyTorch means that the decision tree compiler has no control over lower level details like shared memory, thread synchronization, etc. Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized. Systems like TorchDynamo and TorchInductor do not provide this level of control to the PyTorch program.

These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap.

Rev#B: Are your compiler changes 'portable' across platforms?
The scheduling language, high-level IR and mid-level IR are fully shared. The same scheduling language is used to specify the structure of inference code on CPUs and GPUs. When a schedule doesn't mark any loops as targetting a GPU dimension, CPU code is generated. The high-level and mid-level IR and optimizations on them are fully shared between the CPU and GPU compilation pipelines. The inference computation is represented as loops in both these IRs and therefore the, the representations and transformations on them are common. The compilation pipelines diverge starting starting at the lowering of MIR to low-level IR. Even so, we are able to share significant amount of code between the CPU and GPU lowering due to how in-memory representations are abstracted in SilvanForge (Section 6). For CPUs, exploring only a few configurations is sufficient to find the best schedule as most optimizations are almost always beneficial. Therefore, a heuristic is not needed.

Rev#C: Have people moved on from decision tree models?
As we discuss in the paper, decision trees are widely used in practice. Also, the fact that their popularity has not declined over the last few years is evident from the last few Kaggle surveys. We believe that the retargetable performance that SilvanForge provides will be valuable to the large number of practitioners who use decision trees.

Rev#C: Support for reductions.
MLIR currently supports parallel reductions for value types (scf.reduce) and reducing a dimension of a tensor (linalg.reduce). Neither of these meets the requirements for generating code for decision tree inference. With scf.reduce, every accumulation creates a logical copy and eliminating these copies correctly is a non-trivial problem. With linalg.reduce, the reduction cannot be fused with another loop as the entire tensor of values that are to be reduced need to be written to memory first.

The reduction support we implement in SilvanForge is fairly general. Only the process of identifying reduction loops (Section 5) is specific to decision tree inference (tree loops are identified as reduction loops). All lowering of the ops in the reduction dialect are general and can be used for any other application that requires reductions. 
-------------------------------------------------------------------------

### 2024-07-26 docs/SOSP 2024/rebuttal.txt

Q1: Novelty of scheduling language.

Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees. One of the paper's key contributions lies in recognizing the existence of a scheduling space for this domain and demonstrating the significant performance variations across different schedules. The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like tree-walk interleaving, tree-walk unrolling, reduction optimizations etc. These cannot be supported with existing scheduling languages. This has parallels to how TVM extends Halide with domain-specific primitives for DNN compilation like tensorization and cooperative data loading. Despite the fact that TVM and Halide operate on similar domains (deep learning vs. image processing), such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.

@@

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding the required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work has proposed a small fixed set of optimization techniques, and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

Q2: Comparison with Hummingbird

Hummingbird translates decision tree inference into a series of tensor operations to use PyTorch. While this enables re-targetability, the tensor abstraction is insufficient to express all SilvanForge optimizations. The generated tensor code performs all tree walks in parallel, stores the individual tree predictions in a tensor and finally performs a reduction. Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks). This, as discussed in our paper, is not always useful (Section 2). However. it is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.

Using tensor operations to represent tree traversals obscures the computation's semantics, making it difficult for the tensor compiler to perform these optimizations. Consider the example of walk-interleaving. SilvanForge’s tree-walk interleaving relies on unrolling only the trees whose walks are being interleaved to the same depth. It also reorders trees so that trees with the same depth are processed together. While it maybe possible to extend tensor compilers to perform interleaving, it will be very hard to mimic what SilvanForge does. Current tensor compilers optimize dense matrix computation and do not have the optimizations necessary to interleave instructions across iterations. Even if they did, picking the right tree-walks (same tree different rows, different trees same row etc.) to interleave would be a non-trivial problem.

The tensor compiler cannot perform the required caching and reduction optimizations either. Since the generated tensor code relies on gathering individual features when required, the indirect accesses to trees and rows complicates the analysis of memory access and reuse. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does. Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and one for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.

### 2024-07-26 docs/SOSP 2024/rebuttal.txt

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding the required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work has proposed a small fixed set of optimization techniques, and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

Q2: Comparison with Hummingbird

Hummingbird translates decision tree inference into a series of tensor operations to use PyTorch. While this enables re-targetability, the tensor abstraction is insufficient to express all SilvanForge optimizations. The generated tensor code performs all tree walks in parallel, stores the individual tree predictions in a tensor and finally performs a reduction. Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks). This, as discussed in our paper, is not always useful (Section 2). However. it is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.

Further, using tensor operations to represent tree traversals obscures the computation's semantics, making it difficult for the tensor compiler to perform these optimizations. Consider the example of walk-interleaving. SilvanForge’s tree-walk interleaving relies on unrolling only the trees whose walks are being interleaved to the same depth. It also reorders trees so that trees with the same depth are processed together. While it maybe possible to extend tensor compilers to perform interleaving, it will be very hard to mimic what SilvanForge does. Current tensor compilers optimize dense matrix computation and do not have the optimizations necessary to interleave instructions across iterations. Even if this support is added, picking the right tree-walks (same tree different rows, different trees same row etc.) to interleave would be a non-trivial problem.

@@

The tensor compiler cannot perform the required caching and reduction optimizations either. Since the generated tensor code relies on gathering individual features when required, the indirect accesses to trees and rows complicates the analysis of memory access and reuse. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does. Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and one for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.

The above discussion brings up a more practical system level problem. Expressing the tree inference using PyTorch means that the decision tree compiler has no control over lower level details like shared memory, thread synchronization, etc. Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized. Systems like TorchDynamo and TorchInductor do not provide this level of control to the PyTorch program.

These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap.

Rev#B: Are your compiler changes 'portable' across platforms?
The scheduling language, high-level IR and mid-level IR and optimizations on them are fully shared. The same scheduling language is used to specify the structure of inference code on CPUs and GPUs. When a schedule doesn't mark any loops as targetting a GPU dimension, CPU code is generated. The compilation pipelines diverge starting starting at the lowering of MIR to low-level IR. Even so, we are able to share significant amount of code between the CPU and GPU lowering because of how in-memory representations are abstracted (Section 6).

Rev#C: Support for reductions.
MLIR currently supports parallel reductions for value types (scf.reduce) and reducing a dimension of a tensor (linalg.reduce). Neither of these meets the requirements for generating code for decision tree inference. With scf.reduce, every accumulation creates a logical copy and eliminating these copies correctly is a non-trivial problem. With linalg.reduce, the reduction cannot be fused with another loop as the entire tensor of values that are to be reduced need to be written to memory first. The reduction support we implement in SilvanForge is fairly general. Only the process of identifying reduction loops is currently specific to decision tree inference. All lowering of the ops in the reduction dialect are general.

-------------------------------------------------------------------------
